---
title: "Descriptive statistics"
subtitle: 5mhC expression in kidney cancer
output: 
  html_document: 
    keep_md: yes
---

```{r Global, echo = FALSE, warning = FALSE, message = FALSE}
# Establishing global options
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "markdown")
# Loading libraries and scripts
library(ggplot2)
library(dplyr)
library(simpleR)
library(stringr)
# Personalized ggplot2 theme
gtheme <- theme(
  plot.title = element_text(size = 20, vjust = 1.5),
  axis.title.x = element_text(vjust = -.5, size = 16),
  axis.text.x = element_text(color = "black"),
  axis.title.y = element_text(vjust = 1.5, size = 16),
  axis.text.y = element_text(color = "black"),
  legend.position = "bottom"
)
```

```{r Data}
# Loading the dataset
Data_full <- read.csv("DatasetKidney.csv")
# Fixing column names
colnames(Data_full) <- colnames(Data_full) %>%
  str_replace_all("\\.", "_")
# Selecting clinicopathologic features and 5hmC expression
Data_full <- Data_full %>%
  select(site:fu_months, starts_with("hmc"))
```

In this section we will describe the dataset, including patient's clinical features, preoperative and surgical data, pathologic features of the primary and metastatic tumors, and patient's outcome. For this section we will use the full dataset (N = `r nrow(Data_full)`), analyzing primary and metastatic tumors separately.

```{r Site}
# site
var <- Data_full$site
ggplot(Data_full, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by tumor site") +
  xlab("") +
  ylab("No. patients") +
  scale_y_continuous(limits = c(0, 70), breaks = seq(0, 70, 10)) +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
```

***

## Primary tumors
```{r Primary}
# Selecting only primary tumors
Data <- Data_full %>% filter(site == "Primary")
# age
var <- Data$age
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 5) +
  ggtitle("Distribution of patients by age") +
  xlab("Age, years") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# sex
var <- Data$sex
ggplot(Data, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by sex") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# race
var <- Data$race
var <- reorder(var, -table(var)[var])
ggplot(Data, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by race") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# procedure
var <- Data$procedure
ggplot(Data, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by surgical procedure") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# size
var <- Data$size
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 1.5) +
  ggtitle("Distribution of patients by tumor size") +
  xlab("Tumor size, cm") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# stage
Data_nona <- Data %>% filter(!is.na(stage))
var <- Data_nona$stage
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by tumor stage") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# fuhrman
Data_nona <- Data %>% filter(!is.na(fuhrman))
var <- factor(Data_nona$fuhrman,
  labels = c("Grade 1", "Grade 2", "Grade 3", "Grade 4"))
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by tumor's Fuhrman grade") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# multifocal
Data_nona <- Data %>% filter(!is.na(multifocal))
var <- Data_nona$multifocal
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by tumor multifocality") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# free_margin
Data_nona <- Data %>% filter(!is.na(free_margin))
var <- Data_nona$free_margin
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by tumor invasion of \nsurgical margins") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# other_cancer
Data_nona <- Data %>% filter(!is.na(other_cancer))
var <- Data_nona$other_cancer
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by concurrent malignancies") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# fu_months
var <- Data$fu_months
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 6) +
  ggtitle("Distribution of patients by follow-up length") +
  xlab("Follow-up, months") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# progression
Data_nona <- Data %>% filter(!is.na(progression))
var <- Data_nona$progression
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by tumor progression") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# death
Data_nona <- Data %>% filter(!is.na(death))
var <- Data_nona$death
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by overall mortality") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# dod
Data_nona <- Data %>% filter(!is.na(dod))
var <- Data_nona$dod
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by cancer-related mortality") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# hmc_mean_normal
var <- Data$hmc_mean_normal
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 25) +
  ggtitle("Distribution of mean H-scores in nontumor tissue") +
  xlab("Mean H-score") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# hmc_med_normal
var <- Data$hmc_med_normal
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 25) +
  ggtitle("Distribution of median H-scores in nontumor tissue") +
  xlab("Median H-score") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# hmc_mean_tumor
var <- Data$hmc_mean_tumor
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 20) +
  ggtitle("Distribution of mean H-scores in tumor tissue") +
  xlab("Mean H-score") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# hmc_med_tumor
var <- Data$hmc_med_tumor
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 25) +
  ggtitle("Distribution of median H-scores in tumor tissue") +
  xlab("Mean H-score") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
```

***

## Pathologic features of metastatic tumors
```{r Metastatic}
# Selecting only metastatic tumors
Data <- Data_full %>% filter(site == "Metastatic")
# age
var <- Data$age
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 5) +
  ggtitle("Distribution of patients by age") +
  xlab("Age, years") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# sex
var <- Data$sex
ggplot(Data, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by sex") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# race
var <- Data$race
var <- reorder(var, -table(var)[var])
ggplot(Data, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by race") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# fu_months
var <- Data$fu_months
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 24) +
  ggtitle("Distribution of patients by follow-up length") +
  xlab("Follow-up, months") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# death
Data_nona <- Data %>% filter(!is.na(death))
var <- Data_nona$death
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by overall mortality") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# dod
Data_nona <- Data %>% filter(!is.na(dod))
var <- Data_nona$dod
ggplot(Data_nona, aes(x = var, fill = var)) +
  geom_bar(color = "black") +
  ggtitle("Distribution of patients by cancer-related mortality") +
  xlab("") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.categorical(var)
# hmc_mean_tumor
var <- Data$hmc_mean_tumor
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 20) +
  ggtitle("Distribution of mean H-scores in tumor tissue") +
  xlab("Mean H-score") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
# hmc_med_tumor
var <- Data$hmc_med_tumor
ggplot(Data, aes(x = var, fill = factor(0))) +
  geom_histogram(color = "black", binwidth = 25) +
  ggtitle("Distribution of median H-scores in tumor tissue") +
  xlab("Mean H-score") +
  ylab("No. patients") +
  gtheme + theme(legend.position = "none")
descriptive.numerical(var)
```

