---
title: Expression and outcome in metastatic tumors
subtitle: 5mhC expression in kidney cancer
output: 
  html_document: 
    keep_md: yes
---

```{r Global, echo = FALSE, warning = FALSE, message = FALSE}
# Establishing global options
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "markdown")
# Loading libraries and scripts
library(ggplot2)
library(dplyr)
library(tidyr)
library(simpleR)
library(stringr)
library(grid)
# Personalized ggplot2 theme
gtheme <- theme(
  plot.title = element_text(size = 20, vjust = 1.5),
  axis.title.x = element_text(vjust = -.5, size = 16),
  axis.text.x = element_text(color = "black"),
  axis.title.y = element_text(vjust = 1.5, size = 16),
  axis.text.y = element_text(color = "black"),
  plot.margin = unit(c(3, 1, .5, .5), "lines")
)
```

```{r Plotting}
# Barplots
G_barplot <- function(){
  Z <- max(table(X, Y)) + 4
  W <- (nlevels(X)/2) + .5
  P <- format(fisher.test(X, Y)$p.value, digits = 2, scientific = TRUE)
  ggplot(Data_G, aes(x = X, fill = Y)) +
    geom_bar(color = "black", position = position_dodge()) +
    ylab("No. patients") +
    scale_fill_discrete(name = "5hmC level") +
    gtheme +
    xlab("") +
    scale_y_continuous(limits = c(0, Z), breaks = seq(0, Z, 10)) +
    annotate("text", label = paste("P value =", P), x = W, y = Z) +
    ggtitle(M) +
    gtheme
}
# Table of factor by group
T_Factor <- function(){
  Data_G <- na.omit(Data[,c(V, O)])
  X <- Data_G[,2]
  Y <- Data_G[,1]
  descriptive.categorical.group(X, Y)
}
```

For this section we will use 3 different approaches for evaluating the association of 5hmC and outcome: 1) using 5hmC H-scores; 2) categorizing 5hmC median expression in low/high levels using the median of the distribution as the cutoff point; 3) categorizing 5hmC median expression in low/high levels using the upper tertile of the distribution as the cutoff point.

Outcomes will include overall mortality and cancer-related mortality. This section includes only the analysis of metastatic tumors. Primary tumors are analyzed separately.

```{r Data}
# Load the dataset
Data_full <- read.csv("DatasetKidney.csv")
# Fix column names
colnames(Data_full) <- colnames(Data_full) %>%
  str_replace_all("\\.", "_")
# Selecte clinicopathologic features and 5hmC expression
Data_full <- Data_full %>%
  select(site:fu_months, hmc_med_tumor)
# Recoding fuhrman as factor
Data_full$fuhrman <- factor(Data_full$fuhrman, labels = c(
  "Grade 1", "Grade 2", "Grade 3", "Grade 4"
))
# Create categories of 5hmC expression
Data_full <- Data_full %>%
  mutate(p50 = factor(
    ifelse(hmc_med_tumor > median(hmc_med_tumor),
      "High expression", "Low expression")
  )) %>%
  mutate(q3 = factor(
    ifelse(hmc_med_tumor > quantile(hmc_med_tumor, .66),
      "High expression", "Low expression")
  ))
# Create age and size group categories
Data_full <- Data_full %>%
  mutate(
    age_group = factor(
      ifelse(age > median(age, na.rm = TRUE),
        "Above median age", "Below median age")
    ),
    size_group = factor(
      ifelse(size > median(size, na.rm = TRUE),
        "Above median size", "Below median size")
    )
  )
# Select only metastatic tumors
Data <- Data_full %>%
  filter(site == "Metastatic")
Data <- droplevels(Data)
```

## Approach #1
In this approach we will compare 5hmC `H-scores` against clinicopathologic features and outcome. Comparisons will be made using the Mann-Whitney U test or the Kruskal-Wallis test. Age and tumor size will be categorized in 2 levels using the median age as the cutoff point.

```{r Approach_1}
Y <- Data$hmc_med_tumor
# age_group
X <- Data$age_group
Z <- 200
P <- format(kruskal.test(Y ~ X)$p.value, digits = 2, scientific = TRUE)
ggplot(Data, aes(x = X, y = Y, fill = X)) +
  stat_boxplot(geom = "errorbar", stat_params = list(width = 0.5)) +
  geom_boxplot(color = "black", outlier.shape = 1) +
  ggtitle("Comparison of 5hmC H-scores by \nage group") +
  xlab("Age group") +
  ylab("5hmC H-score") +
  scale_y_continuous(limits = c(0, Z), breaks = seq(0, Z, 50)) +
  annotate("text", label = paste("P value =", P),
    x = 1.5, y = Z) +
  gtheme + theme(legend.position = "none")
descriptive.numerical.group(Y, X)
# sex
X <- Data$sex
Z <- 200
P <- format(kruskal.test(Y ~ X)$p.value, digits = 2, scientific = TRUE)
ggplot(Data, aes(x = X, y = Y, fill = X)) +
  stat_boxplot(geom = "errorbar", stat_params = list(width = 0.5)) +
  geom_boxplot(color = "black", outlier.shape = 1) +
  ggtitle("Comparison of 5hmC H-scores by \nsex") +
  xlab("Sex") +
  ylab("5hmC H-score") +
  scale_y_continuous(limits = c(0, Z), breaks = seq(0, Z, 50)) +
  annotate("text", label = paste("P value =", P),
    x = 1.5, y = Z) +
  gtheme + theme(legend.position = "none")
descriptive.numerical.group(Y, X)
# race
X <- Data$race
Z <- 200
P <- format(kruskal.test(Y ~ X)$p.value, digits = 2, scientific = TRUE)
ggplot(Data, aes(x = X, y = Y, fill = X)) +
  stat_boxplot(geom = "errorbar", stat_params = list(width = 0.5)) +
  geom_boxplot(color = "black", outlier.shape = 1) +
  ggtitle("Comparison of 5hmC H-scores by \nrace") +
  xlab("Race") +
  ylab("5hmC H-score") +
  scale_y_continuous(limits = c(0, Z), breaks = seq(0, Z, 50)) +
  annotate("text", label = paste("P value =", P),
    x = 1.5, y = Z) +
  gtheme + theme(legend.position = "none")
descriptive.numerical.group(Y, X)
# size_group
# stage
# fuhrman
# multifocal
# free_margin
# progression
# death
Data_nona <- Data %>% filter(!is.na(death))
X <- Data_nona$death
Y <- Data_nona$hmc_med_tumor
Z <- 200
P <- format(kruskal.test(Y ~ X)$p.value, digits = 2, scientific = TRUE)
ggplot(Data_nona, aes(x = X, y = Y, fill = X)) +
  stat_boxplot(geom = "errorbar", stat_params = list(width = 0.5)) +
  geom_boxplot(color = "black", outlier.shape = 1) +
  ggtitle("Comparison of 5hmC H-scores by \noverall mortality") +
  xlab("Overall mortality") +
  ylab("5hmC H-score") +
  scale_y_continuous(limits = c(0, Z), breaks = seq(0, Z, 50)) +
  annotate("text", label = paste("P value =", P),
    x = 1.5, y = Z) +
  gtheme + theme(legend.position = "none")
descriptive.numerical.group(Y, X)
# dod
Data_nona <- Data %>% filter(!is.na(dod))
X <- Data_nona$dod
Y <- Data_nona$hmc_med_tumor
Z <- 200
P <- format(kruskal.test(Y ~ X)$p.value, digits = 2, scientific = TRUE)
ggplot(Data_nona, aes(x = X, y = Y, fill = X)) +
  stat_boxplot(geom = "errorbar", stat_params = list(width = 0.5)) +
  geom_boxplot(color = "black", outlier.shape = 1) +
  ggtitle("Comparison of 5hmC H-scores by \ncancer-related mortality") +
  xlab("Cancer-related mortality") +
  ylab("5hmC H-score") +
  scale_y_continuous(limits = c(0, Z), breaks = seq(0, Z, 50)) +
  annotate("text", label = paste("P value =", P),
    x = 1.5, y = Z) +
  gtheme + theme(legend.position = "none")
descriptive.numerical.group(Y, X)
```

***

## Approach # 2

In this approach we will compare 5hmC levels (categorizing 5hmC expression in low/high levels using the `median` as the cutoff point) against clinicopathologic features and outcome. Comparisons will be made using the Fisher's exact test.

```{r Approach_2}
V <- "p50"
# age_group
O <- "age_group"
M <- "Comparison of 5hmC levels \nby age"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# sex
O <- "sex"
M <- "Comparison of 5hmC levels \nby pT sex"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# race
O <- "race"
M <- "Comparison of 5hmC levels \nby race"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# size_group
# stage
# fuhrman
# multifocal
# free_margin
# progression
# death
O <- "death"
M <- "Comparison of 5hmC levels \nby overall mortality"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# dod
O <- "dod"
M <- "Comparison of 5hmC levels \nby cancer-related mortality"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
```

***

## Approach # 3

In this approach we will compare 5hmC levels (categorizing 5hmC expression in low/high levels using the `upper tertile` as the cutoff point) against clinicopathologic features and outcome. Comparisons will be made using the Fisher's exact test.

```{r Approach_3}
V <- "q3"
# age_group
O <- "age_group"
M <- "Comparison of 5hmC levels \nby age"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# sex
O <- "sex"
M <- "Comparison of 5hmC levels \nby pT sex"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# race
O <- "race"
M <- "Comparison of 5hmC levels \nby race"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# size_group
# stage
# fuhrman
# multifocal
# free_margin
# progression
# death
O <- "death"
M <- "Comparison of 5hmC levels \nby overall mortality"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
# dod
O <- "dod"
M <- "Comparison of 5hmC levels \nby cancer-related mortality"
Data_G <- na.omit(Data[,c(V, O)]); X <- Data_G[,2]; Y <- Data_G[,1]
G_barplot(); T_Factor()
```

