---
title: Expression and survival in primary tumors
subtitle: 5mhC expression in kidney cancer
output: 
  html_document: 
    keep_md: yes
---

```{r Global, echo = FALSE, warning = FALSE, message = FALSE}
# Establishing global options
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "markdown")
# Loading libraries and scripts
library(ggplot2)
library(ggfortify)
library(survival)
library(dplyr)
library(tidyr)
library(stringr)
library(grid)
# Personalized ggplot2 theme
gtheme <- theme(
  plot.title = element_text(size = 20, vjust = 1.5),
  axis.title.x = element_text(vjust = 0, size = 16),
  axis.text.x = element_text(color = "black"),
  axis.title.y = element_text(vjust = 1.5, size = 16),
  axis.text.y = element_text(color = "black"),
  legend.position = c(.9, .9),
  legend.title = element_blank(),
  plot.margin = unit(c(1.5, 1, 1.5, 1), "lines")
)
```

```{r gsurv}
# ggplot2 survival curves using ggfortify
gsurv <- function(FU, outcome, status, Y = 0.1){
  ht <- survdiff(Surv(FU, as.numeric(outcome)) ~ status)
  p <- format(pchisq(ht$chisq, df = 1, lower.tail = FALSE), digits = 2)
  fit <- survfit(Surv(FU, as.numeric(outcome)) ~ status)
  autoplot(fit, conf.int = FALSE, censor = FALSE, surv.size = 1) +
    ylab("Survival") +
    xlab("Follow-up, months") +
    annotate("text", label = paste("Log-rank P value =", p), x = 12, y = Y) +
    scale_colour_discrete(labels = levels(status)) +
    scale_x_continuous(limits = c(0, 60), breaks = seq(0, 60, 12)) +
    gtheme
}
```

In this section we will plot survival curves for overall mortality and cancer-related mortality according to clinicopathologic features and 5hmC expression, categorizing 5hmC expression in low/high levels using as cutoff points: 1) the median; 2) the upper tertile. Follow-up length will be defined as months from nephrectomy.

For age and tumor size we will use the median to define categorical groups. Fuhrman grades and pT stages will also be colapsed in 2-level categorical groups.

This section includes only the analysis of primary tumors. Metastatic tumors are analyzed separately.

```{r Data}
# Load the dataset
Data_full <- read.csv("DatasetKidney.csv")
# Fix column names
colnames(Data_full) <- colnames(Data_full) %>%
  str_replace_all("\\.", "_")
# Selecte clinicopathologic features and 5hmC expression
Data_full <- Data_full %>%
  select(site:fu_months, hmc_med_tumor)
# Recoding fuhrman as factor
Data_full$fuhrman <- factor(Data_full$fuhrman, labels = c(
  "Grade 1", "Grade 2", "Grade 3", "Grade 4"
))
# Create categories of 5hmC expression
Data_full <- Data_full %>%
  mutate(p50 = factor(
    ifelse(hmc_med_tumor > median(hmc_med_tumor),
      "High expression", "Low expression")
  )) %>%
  mutate(q3 = factor(
    ifelse(hmc_med_tumor > quantile(hmc_med_tumor, .66),
      "High expression", "Low expression")
  ))
# Create age and size group categories
Data_full <- Data_full %>%
  mutate(
    age_group = factor(
      ifelse(age > median(age, na.rm = TRUE),
        "Above median age", "Below median age")
    ),
    size_group = factor(
      ifelse(size > median(size, na.rm = TRUE),
        "Above median size", "Below median size")
    )
  )
# Create Fuhrman grades category (> Grade 2)
Data_full <- Data_full %>%
  mutate(
    fuhrman_high = factor(
      ifelse(as.numeric(fuhrman) > 2, "High grade", "Low grade")
    )
  )
# Create pT stages category (> pT1a)
Data_full <- Data_full %>%
  mutate(
    stage_high = factor(
      ifelse(as.numeric(stage) > 1, "High pT", "Low pT")
    )
  )
# Select only primary tumors
Data <- Data_full %>%
  filter(site == "Primary")
Data <- droplevels(Data)
# Defining follow-up
FU <- Data$fu_months
```

***

## Clinicopathologic features and survival
In this section we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality considering clinicopathologic features. We will compare survival curves using the `log-rank test`.

### Overall mortality
```{r Clinicopath_Death}
# Establishing outcome
outcome <- Data$death
# Patient's age
status <- Data$age_group
gsurv(FU, outcome, status, Y = 0.625) +
  ggtitle("Patient's age and overall mortality") +
  theme(legend.position = c(0.86, 0.9))
# Patient's sex
status <- Data$sex
gsurv(FU, outcome, status, Y = 0.65) +
  ggtitle("Patient's sex and overall mortality") +
  theme(legend.position = c(0.92, 0.9))
# size_group
status <- Data$size_group
gsurv(FU, outcome, status, Y = 0.45) +
  ggtitle("Tumor size and overall mortality") +
  theme(legend.position = c(0.85, 0.8))
# stage_high
status <- Data$stage_high
gsurv(FU, outcome, status, Y = 0.45) +
  ggtitle("High tumor stage (>pT1a) and overall mortality") +
  theme(legend.position = c(0.9, 0.85))
# fuhrman_high
status <- Data$fuhrman_high
gsurv(FU, outcome, status, Y = 0.45) +
  ggtitle("High Fuhrman grade (>2) and overall mortality") +
  theme(legend.position = c(0.9, 0.75))
# multifocal
status <- Data$multifocal
gsurv(FU, outcome, status, Y = 0.65) +
  ggtitle("Tumor multifocality and overall mortality") +
  theme(legend.position = c(0.925, 0.9))
# free_margin
status <- Data$free_margin
gsurv(FU, outcome, status, Y = 0.65) +
  ggtitle("Invasion of surgical margins and overall mortality") +
  theme(legend.position = c(0.925, 0.9))
# progression
status <- Data$progression
gsurv(FU, outcome, status, Y = 0.05) +
  ggtitle("Tumor progression and overall mortality") +
  theme(legend.position = c(0.925, 0.9))
```

***

### Cancer-related mortality
```{r Clinicopath_DOD}
# Establishing outcome
outcome <- Data$dod
# Patient's age
status <- Data$age_group
gsurv(FU, outcome, status, Y = 0.85) +
  ggtitle("Patient's age and cancer-related mortality") +
  theme(legend.position = c(0.86, 0.8))
# Patient's sex
status <- Data$sex
gsurv(FU, outcome, status, Y = 0.80) +
  ggtitle("Patient's sex and cancer-related mortality") +
  theme(legend.position = c(0.92, 0.85))
# size_group
status <- Data$size_group
gsurv(FU, outcome, status, Y = 0.80) +
  ggtitle("Tumor size and cancer-related mortality") +
  theme(legend.position = c(0.85, 0.8))
# stage_high
status <- Data$stage_high
gsurv(FU, outcome, status, Y = 0.80) +
  ggtitle("High tumor stage (>pT1a) and \ncancer-related mortality") +
  theme(legend.position = c(0.9, 0.85))
# fuhrman_high
status <- Data$fuhrman_high
gsurv(FU, outcome, status, Y = 0.75) +
  ggtitle("High Fuhrman grade (>2) and \ncancer-related mortality") +
  theme(legend.position = c(0.9, 0.75))
# multifocal
status <- Data$multifocal
gsurv(FU, outcome, status, Y = 0.85) +
  ggtitle("Tumor multifocality and \ncancer-related mortality") +
  theme(legend.position = c(0.925, 0.8))
# free_margin
status <- Data$free_margin
gsurv(FU, outcome, status, Y = 0.91) +
  ggtitle("Invasion of surgical margins and \ncancer-related mortality") +
  theme(legend.position = c(0.925, 0.875))
# progression
status <- Data$progression
gsurv(FU, outcome, status, Y = 0.05) +
  ggtitle("Tumor progression and \ncancer-related mortality") +
  theme(legend.position = c(0.925, 0.825))
```

## 5hmC levels and survival: Approach #1
In this approch we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality using the `median` H-score as the cutoff point for defining low/high 5hmC levels. We will compare survival curves using the `log-rank test`.

```{r Approach1}
# 5hmC and overall mortality
outcome <- Data$death
status <- Data$p50
gsurv(FU, outcome, status, Y = 0.63) +
  ggtitle("5hmC expression levels and \noverall mortality") +
  theme(legend.position = c(0.875, 0.875))
# 5hmC and cancer-related mortality
outcome <- Data$dod
gsurv(FU, outcome, status, Y = 0.865) +
  ggtitle("5hmC expression levels and \ncancer-related mortality") +
  theme(legend.position = c(0.875, 0.8))
```

***

## 5hmC levels and survival: Approach #2
In this approch we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality using the upper tertile `Q3` H-score as the cutoff point for defining low/high 5hmC levels. We will compare survival curves using the `log-rank test`.

```{r Approach2}
# 5hmC and overall mortality
outcome <- Data$death
status <- Data$q3
gsurv(FU, outcome, status, Y = 0.72) +
  ggtitle("5hmC expression levels and \noverall mortality") +
  theme(legend.position = c(0.875, 0.875))
# 5hmC and cancer-related mortality
outcome <- Data$dod
gsurv(FU, outcome, status, Y = 0.9) +
  ggtitle("5hmC expression levels and \ncancer-related mortality") +
  theme(legend.position = c(0.875, 0.8))
```

***
