---
title: Expression and survival in metastatic tumors
subtitle: 5mhC expression in kidney cancer
output: 
  html_document: 
    keep_md: yes
---

```{r Global, echo = FALSE, warning = FALSE, message = FALSE}
# Establishing global options
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "markdown")
# Loading libraries and scripts
library(ggplot2)
library(ggfortify)
library(survival)
library(dplyr)
library(tidyr)
library(stringr)
library(grid)
# Personalized ggplot2 theme
gtheme <- theme(
  plot.title = element_text(size = 20, vjust = 1.5),
  axis.title.x = element_text(vjust = 0, size = 16),
  axis.text.x = element_text(color = "black"),
  axis.title.y = element_text(vjust = 1.5, size = 16),
  axis.text.y = element_text(color = "black"),
  legend.position = c(.9, .9),
  legend.title = element_blank(),
  plot.margin = unit(c(1.5, 1, 1.5, 1), "lines")
)
```

```{r gsurv}
# ggplot2 survival curves using ggfortify
gsurv <- function(FU, outcome, status, Y = 0.1){
  ht <- survdiff(Surv(FU, as.numeric(outcome)) ~ status)
  p <- format(pchisq(ht$chisq, df = 1, lower.tail = FALSE), digits = 2)
  fit <- survfit(Surv(FU, as.numeric(outcome)) ~ status)
  autoplot(fit, conf.int = FALSE, censor = FALSE, surv.size = 1) +
    ylab("Survival") +
    xlab("Follow-up, months") +
    annotate("text", label = paste("Log-rank P value =", p), x = 48, y = Y) +
    scale_colour_discrete(labels = levels(status)) +
    scale_x_continuous(limits = c(0, 240), breaks = seq(0, 240, 24)) +
    gtheme
}
```

In this section we will plot survival curves for cancer-related mortality  (since all deaths are cancer-related in this group) according to clinicopathologic features and 5hmC expression, categorizing 5hmC expression in low/high levels using as cutoff points: 1) the median; 2) the upper tertile. Follow-up length will be defined as months from nephrectomy.

For age we will use the median to define categorical groups.

This section includes only the analysis of metastatic tumors. Primary tumors are analyzed separately.

```{r Data}
# Load the dataset
Data_full <- read.csv("DatasetKidney.csv")
# Fix column names
colnames(Data_full) <- colnames(Data_full) %>%
  str_replace_all("\\.", "_")
# Selecte clinicopathologic features and 5hmC expression
Data_full <- Data_full %>%
  select(site:fu_months, hmc_med_tumor)
# Recoding fuhrman as factor
Data_full$fuhrman <- factor(Data_full$fuhrman, labels = c(
  "Grade 1", "Grade 2", "Grade 3", "Grade 4"
))
# Create categories of 5hmC expression
Data_full <- Data_full %>%
  mutate(p50 = factor(
    ifelse(hmc_med_tumor > median(hmc_med_tumor),
      "High expression", "Low expression")
  )) %>%
  mutate(q3 = factor(
    ifelse(hmc_med_tumor > quantile(hmc_med_tumor, .66),
      "High expression", "Low expression")
  ))
# Create age group category
Data_full <- Data_full %>%
  mutate(
    age_group = factor(
      ifelse(age > median(age, na.rm = TRUE),
        "Above median age", "Below median age")
    )
  )

# Select only metastatic tumors
Data <- Data_full %>%
  filter(site == "Metastatic")
Data <- droplevels(Data)
# Defining follow-up
FU <- Data$fu_months
```

***

## Clinical features and survival
In this section we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality considering clinical features. We will compare survival curves using the `log-rank test`.

### Cancer-related mortality
```{r Clinicopath_DOD}
# Establishing outcome
outcome <- Data$dod
# Patient's age
status <- Data$age_group
gsurv(FU, outcome, status, Y = 0.25) +
  ggtitle("Patient's age and cancer-related mortality") +
  theme(legend.position = c(0.85, 0.875))
# Patient's sex
status <- Data$sex
gsurv(FU, outcome, status, Y = 0.20) +
  ggtitle("Patient's sex and cancer-related mortality") +
  theme(legend.position = c(0.9, 0.875))
# size_group
# stage_high
# fuhrman_high
# multifocal
# free_margin
# progression
```

## 5hmC levels and survival: Approach #1
In this approch we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality using the `median` H-score as the cutoff point for defining low/high 5hmC levels. We will compare survival curves using the `log-rank test`.

```{r Approach1}
# 5hmC and cancer-related mortality
outcome <- Data$dod
status <- Data$p50
gsurv(FU, outcome, status, Y = 0.05) +
  ggtitle("5hmC expression levels and \ncancer-related mortality") +
  theme(legend.position = c(0.875, 0.875))
```

***

## 5hmC levels and survival: Approach #2
In this approch we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality using the upper tertile `Q3` H-score as the cutoff point for defining low/high 5hmC levels. We will compare survival curves using the `log-rank test`.

```{r Approach2}
# 5hmC and cancer-related mortality
outcome <- Data$dod
status <- Data$q3
gsurv(FU, outcome, status, Y = 0.05) +
  ggtitle("5hmC expression levels and \ncancer-related mortality") +
  theme(legend.position = c(0.875, 0.875))
```

***
