---
title: Expression and regression models in metastatic tumors
subtitle: 5mhC expression in kidney cancer
output: 
  html_document: 
    keep_md: yes
---

```{r Global, echo = FALSE, warning = FALSE, message = FALSE}
# Establishing global options
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "markdown")
# Loading libraries and scripts
library(survival)
library(dplyr)
library(tidyr)
library(stringr)
```

```{r Regression}
# Logistic regression tables
logistic.table <- function(outcome, predictors, varlabels){
  model <- list()
  OR.center <- list()
  OR.low <- list()
  OR.high <- list()
  OR.p <- list()
  for(i in 1:length(predictors)) {
    model[[i]] <- glm(outcome ~ predictors[[i]], family = binomial)
    OR.center[[i]] <-format(exp(coef(model[[i]]))[2], digits = 2, nsmall = 2)
    OR.low[[i]] <- format(exp(confint(model[[i]]))[2, 1], digits = 2, nsmall = 2)
    OR.high[[i]] <- format(exp(confint(model[[i]]))[2, 2], digits = 2, nsmall = 2)
    OR.p[[i]] <- format(summary(model[[i]])$coef[2, 4], digits = 2, width  = 6)
  }
  OR.center <- unlist(OR.center)
  OR.low <- unlist(OR.low)
  OR.high <- unlist(OR.high)
  OR.p <- unlist(OR.p)
  OR.table <- cbind("Variables" = varlabels, "OR" = OR.center,
    "Lower 95% CI" = OR.low, "Upper 95% CI" = OR.high, "P value" = OR.p)
  knitr::kable(OR.table, row.names = FALSE, align = c("l", "c", "c", "c", "c"))
}
# Cox regression tables
cox.table <- function(outcome, fu, predictors, varlabels){
  outcome <- as.numeric(outcome)
  survival.obj <- survival::Surv(fu, outcome)
  model <- list()
  HR.center <- list()
  HR.low <- list()
  HR.high <- list()
  HR.p <- list()
  for(i in 1:length(predictors)) {
    model[[i]] <- survival::coxph(survival.obj ~ predictors[[i]])
    HR.center[[i]] <-format(summary(model[[i]])$conf.int[1], digits = 2, nsmall = 2)
    HR.low[[i]] <- format(summary(model[[i]])$conf.int[3], digits = 2, nsmall = 2)
    HR.high[[i]] <- format(summary(model[[i]])$conf.int[4], digits = 2, nsmall = 2)
    HR.p[[i]] <- format(summary(model[[i]])$logtest[3], digits = 2, width  = 6)
  }
  HR.center <- unlist(HR.center)
  HR.low <- unlist(HR.low)
  HR.high <- unlist(HR.high)
  HR.p <- unlist(HR.p)
  HR.table <- cbind("Variables" = varlabels, "HR" = HR.center,
    "Lower 95% CI" = HR.low, "Upper 95% CI" = HR.high, "P value" = HR.p)
  knitr::kable(HR.table, row.names = FALSE, align = c("l", "c", "c", "c", "c"))
}
```

In this section we will build regression models for cancer-related mortality (since all deaths are cancer-related deaths) according to clinical features and 5hmC expression. 5hmC expression will be categorized in low/high levels using as cutoff points: 1) the median; 2) the upper tertile. 

For age we will use the median to define categorical groups.

We will begin building uncondicional binary logistic regression models to estimate the odds ratios. We then will build Cox's proportional hazards regression models to estimate the hazard ratios. Follow-up length will be defined as months from nephrectomy.

This section includes only the analysis of metastatic tumors. Primary tumors are analyzed separately.

```{r Data}
# Load the dataset
Data_full <- read.csv("DatasetKidney.csv")
# Fix column names
colnames(Data_full) <- colnames(Data_full) %>%
  str_replace_all("\\.", "_")
# Selecte clinicopathologic features and 5hmC expression
Data_full <- Data_full %>%
  select(site:fu_months, hmc_med_tumor)
# Recoding fuhrman as factor
Data_full$fuhrman <- factor(Data_full$fuhrman, labels = c(
  "Grade 1", "Grade 2", "Grade 3", "Grade 4"
))
# Create categories of 5hmC expression
Data_full <- Data_full %>%
  mutate(p50 = factor(
    ifelse(hmc_med_tumor > median(hmc_med_tumor),
      "High expression", "Low expression")
  )) %>%
  mutate(q3 = factor(
    ifelse(hmc_med_tumor > quantile(hmc_med_tumor, .66),
      "High expression", "Low expression")
  ))
# Create age group category
Data_full <- Data_full %>%
  mutate(
    age_group = factor(
      ifelse(age > median(age, na.rm = TRUE),
        "Above median age", "Below median age")
    ),
    size_group = factor(
      ifelse(size > median(size, na.rm = TRUE),
        "Above median size", "Below median size")
    )
  )
# Select only metastatic tumors
Data <- Data_full %>%
  filter(site == "Metastatic")
Data <- droplevels(Data)
# Define follow-up
fu <- Data$fu_months
```

***

## Logistic regression: Cancer-related mortality

```{r Predictors}
predictors <- with(Data, list(
  "Age group: Above vs. below median" = age_group,
  "Sex" = sex,
  "Race" = race,
  "High 5hmC expression (upper median)" = p50,
  "High 5hmC expression (upper tertile)" = q3
))
varlabels <- names(predictors)
```

```{r OR_dod}
outcome <- Data$dod
logistic.table(outcome, predictors, varlabels)
```

***

## Cox regression: Cancer-related mortality

```{r HR_dod}
outcome <- Data$dod
cox.table(outcome, fu, predictors, varlabels)
```

