---
title: Correlation with mTOR markers
subtitle: 5mhC expression in kidney cancer
output: 
  html_document: 
    keep_md: yes
---

```{r Global, echo = FALSE, warning = FALSE, message = FALSE}
# Establishing global options
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "markdown")
# Loading libraries and scripts
library(corrplot)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(grid)
# Personalized ggplot2 theme
gtheme <- theme(
  plot.title = element_text(size = 20, vjust = 1.5),
  axis.title.x = element_text(vjust = 0, size = 16),
  axis.text.x = element_text(color = "black"),
  axis.title.y = element_text(vjust = 1.5, size = 16),
  axis.text.y = element_text(color = "black"),
  legend.position = c(.9, .9),
  legend.title = element_blank(),
  plot.margin = unit(c(2, 1, 2, 1), "lines")
)
```

In this section we will evaluate the correlation between 5hmC median expression and the following markers: PTEN, p-AKT, p-S6, p-4EBP1, p27, c-MYC, and HIF. Data on mTOR markers expression is available only for primary tumors.

```{r Data}
# Load the dataset
Data_full <- read.csv("DatasetKidney.csv")
# Fix column names
colnames(Data_full) <- colnames(Data_full) %>%
  str_replace_all("\\.", "_")
# Select primary tumors and biomarkers
Data <- Data_full %>%
  filter(site == "Primary") %>%
  select(hmc_med_tumor, cmyc_tumor:pten_tumor)
# Listo of biomarkers
biomarkers <- c("hmC", "cMYC", "p27", "pAKT", "pS6", "p4EBP1", "HIF", "PTEN")
colnames(Data) <- biomarkers
```

```{r gcor}
gcor <- function(X = 130) {
  C <- cor.test(VarX, VarY)
  C_P <- format(C$p.value, digits = 2, scientific = TRUE)
  C_C <- format(C$estimate, digits = 2)
  Y <- max(VarY, na.rm = T) - 0.1*max(VarY, na.rm = T)
  ggplot(Data, aes(x = VarX, y = VarY)) +
    geom_point(color = "darkblue", size = 3) +
    geom_smooth(method = lm) +
    annotate("text",
           label = paste("P value =", C_P, "\nCoef =", C_C),
           x = X, y = Y, hjust = 0) +
    xlab("5hmC expression") +
    ylab(Ylab) +
    ggtitle(paste("Correlation of 5hmC and\n", Ylab)) +
    gtheme
}
```


## Correlation matrix
First we will plot a correlation matrix of all biomarkers. Color intensity is related to the values of the correlation coefficients; the actual values of the correlation coefficients are also included. Correlation coefficients were reordered using hierarchical clustering.

```{r Cor_1, fig.width = 9, fig.height = 9}
cor_table <- cor(Data, use = "pairwise.complete.obs")
corrplot(cor_table,
  method = "color",
  diag = FALSE,
  outline = TRUE,
  addCoef.col = "black",
  order = "hclust"
)
```

```{r scatterplots}
# Establish VarX for all scatterplots
VarX <- Data$hmC
# phos-AKT
VarY <- Data$pAKT
Ylab <- "phos-AKT expression"
gcor()
# phos-S6
VarY <- Data$pS6
Ylab <- "phos-S6 expression"
gcor()
# phos-4EBP1
VarY <- Data$p4EBP1
Ylab <- "phos-4EBP1 expression"
gcor()
# c-MYC
VarY <- Data$cMYC
Ylab <- "c-MYC expression"
gcor()
# p27
VarY <- Data$p27
Ylab <- "p27 expression"
gcor()
# p27
VarY <- Data$PTEN
Ylab <- "PTEN expression"
gcor()
# HIF
VarY <- Data$HIF
Ylab <- "HIF expression"
gcor(X = 5)
```

